---
title: "Bird Joins"
author: "Halina Do-Linh"
date: "8/17/2021"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
library(here)
library(tidyverse)
library(lubridate)
library(janitor)
```

# Task One

Read in the data sets and store the data frames as bird_observations, sites, surveys, and taxalist (it should be clear from the raw file names which is which)

```{r}
bird_observations <- read_csv(here::here("data", "52_pp52_birds_1.csv"))

sites <- read_csv(here::here("data", "52_pp52_sites_1.csv"))

surveys <- read_csv(here::here("data", "52_pp52_surveys_1.csv"))

taxalist <- read_csv(here::here("data", "52_pp52_taxalist_1.csv"))
```

Create a subset of bird_observations called birds_subset that only contains observations for birds with species id “BHCO” and “RWBL”, and from sites with site ID “LI-W” and “NU-C”
```{r}
birds_subset <- bird_observations %>% 
  filter(species_id %in% c("BHCO", "RWBL")) %>% 
  filter(site_id %in% c("LI-W", "NU-C"))
```

## Left Join Practice 

Use left join(s) to update birds_subset so that it also includes sites and taxalist information. For each join, include an explicit argument saying which variable you are joining by (even if it will just assume the correct one for you). Store the updated data frame as birds_left. Make sure to look at the output - is what it contains consistent with what you expected it to contain?
```{r}
birds_left <- birds_subset %>% 
  left_join(y = sites, by = "site_id") %>% 
  left_join(y = taxalist, by = "species_id")

# when in doubt be explicit had a weird error come up saying that "species_id" did not exist

# don't see x = birds_subset here because piping in birds_subset means it automatically assumes birds_subset as x.

```

## Full Practice

First, answer: what do you expect a full_join() between birds_subset and sites to contain?

A: I expect to see all columns and all observations from birds_subset and all columns from sites to be merged into one dataset because in a full join everything is kept. 

Write code to full_join the birds_subset and sites data into a new object called birds_full. Explicitly include the variable you’re joining by. Look at the output. Is it what you expected?

```{r}
birds_full <- birds_subset %>% 
  full_join(x = birds_subset, y = sites, by = "site_id")
```

## Inner Join Practice

First, answer: what do you expect an inner_join() between birds_subset and taxalist to contain?

A: I expect to only have data merge together based on "species_id" and since there is only "BHCO" and "RWBL" in birds_subset then I will only retain data related to those two species. I will also expect to see the columns from taxalist - "common_name" and "asu_itis" to be merged into the dataset. 

Write code to inner_join() the birds_subset and taxalist, called birds_inner. Include an argument for what variable you’ll be joining by.

```{r}
birds_inner <- birds_subset %>% 
  inner_join(x = birds_subset, y = taxalist, by = "species_id")
```

What would you get if instead of inner_join() you’d used left_join() for this example? Try it out & check. Why does that make sense for this scenario? It what case would you expect the outcome to differ from an inner_join()?

```{r}
birds_inner_left <- birds_subset %>% 
  left_join(x = birds_subset, y = taxalist, by = "species_id")
```

I have the same dataset after using both inner() and left_join() where x = birds_subset and y = taxalist, by = speices_id. We expect the same here because there is no species_id in birds_subset that doesn't match in taxalist. If there was a species_id in birds_subset that did NOT exist in taxalist, then we would have expected it to see it in the left join because in the left join everything in the left (or x) is kept. If it was an inner join then we would not see the bird species id that did NOT exist in taxalist. It would be removed because inner only keeps what matches in both. 

# Task 2

Starting with your birds object, rename the notes column to bird_obs_notes (so this doesn’t conflict with notes in the surveys dataset)

```{r}
bird_observations <- bird_observations %>% 
  rename(bird_obs_notes = notes)
```

Then, create a subset that contains all observations in the birds dataset, joins the taxonomic, site and survey information to it, and is finally limited to only columns survey_date, common_name, park_name, and bird_count. You can decide the order that you want to create this in (e.g. limit the columns first, then join, or the other way around).

```{r}
bird_observations <- bird_observations %>% 
  full_join(y = taxalist, by = "species_id") %>% 
  full_join(y = sites, by = "site_id") %>% 
  full_join(y = surveys, by = "site_id")
```

